---
title: "BuildMapImages"
author: "Dan Murphy"
date: "August 11, 2016"
output: github_document
---

The goal of the maps shown below is to indicate how the  minimum travel time for a fire engine will change after construction of a new fire station.  

```{r, message=FALSE,warning=FALSE}
library(ggmap)
library(dplyr)
```

First, the minimum travel time is calculated under the existing scenario with three fire stations and then again after construction of the new fire station.  

```{r, eval=FALSE}

LocationsDF$SecondsFromMain <- as.numeric(levels(LocationsDF$SecondsFromMain))[LocationsDF$SecondsFromMain]
LocationsDF$SecondsFromSta2 <- as.numeric(levels(LocationsDF$SecondsFromSta2))[LocationsDF$SecondsFromSta2]
LocationsDF$SecondsFromSta3 <- as.numeric(levels(LocationsDF$SecondsFromSta3))[LocationsDF$SecondsFromSta3]
LocationsDF$SecondsFromSta4 <- as.numeric(levels(LocationsDF$SecondsFromSta4))[LocationsDF$SecondsFromSta4]

LocationsDF <- LocationsDF %>% rowwise() %>% mutate(MinTravelExistingFireStations= 
                                                    min(SecondsFromMain,SecondsFromSta2,SecondsFromSta3))
LocationsDF <- LocationsDF %>% rowwise() %>% mutate(MinTravelw4FireStations = 
                                                    min(SecondsFromMain,SecondsFromSta2,
                                                        SecondsFromSta3,SecondsFromSta4))
```

```{r, echo=FALSE}
load(file="C:/Users/murph/Desktop/R/Firehouse Project/.RData")

LocationsDF <- LocationsDF.Alarms


```

Then, using the ggmap package, qmap pulls the locations image.  With ggplot, locations of alarms are overlayed, which were identified in the GoogleMapsTravelEstimates script with the geocode function

```{r, message=FALSE, warning=FALSE}
CI_Map <- qmap("Central Islip, New York", zoom=13, color="bw", legend="topleft")

MapLocationsDF.Alarms <- LocationsDF %>% filter(MinTravelExistingFireStations<500)

CI_Map + geom_point(aes(x=lon, y=lat, colour=MinTravelExistingFireStations,size=2), 
                    data=MapLocationsDF.Alarms) + 
                    theme(legend.position=c(0,1)) + 
                    labs(colour = "Min Travel Time\nExisting Fire Stations\n(Seconds)") + 
                    guides(size=FALSE)
CI_Map + geom_point(aes(x=lon, y=lat, colour=MinTravelw4FireStations, size=2), 
                    data=MapLocationsDF.Alarms) + 
                    theme(legend.position=c(0,1)) + 
                    labs(colour = "Min Travel Time\nWith 4 Fire Stations\n(Seconds)") + 
                    guides(size=FALSE)

```

This last map shows the change in travel times resulting from the new station.  

```{r,message=FALSE, warning=FALSE}

MapLocationsDF.Alarms$MinTravelChange <- MapLocationsDF.Alarms$MinTravelExistingFireStations - 
                                         MapLocationsDF.Alarms$MinTravelw4FireStations

CI_Map + geom_point(aes(x=lon, y=lat, colour=MinTravelChange,size=2), 
                    data=MapLocationsDF.Alarms) + 
                    theme(legend.position=c(0,1)) + 
                    labs(colour = "Min Travel Time Change\nWith New Fire Station\n(Seconds)") +     
                    guides(size=FALSE)

```
